\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{
An Axiomatic Onto-Epistemic Model \\
of the Hebûn--Zanabûn Framework
}

\müdebbir{
Veysi yê MALA SAF \\
Hebûn–Zanabûn Civilizational Research
}

\date{\today}

\newtheorem{definition}{Definition}
\newtheorem{axiom}{Axiom}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

\begin{document}

\maketitle

\begin{abstract}
We present a formal axiomatic model of the Hebûn--Zanabûn framework,
integrating ontological grounding, epistemological structure,
and ethical calibration within a unified formal system.
The goal is to define a minimal consistent axiomatic core
capable of supporting layered civilizational reasoning.
\end{abstract}

\section{Introduction}

The Hebûn--Zanabûn system proposes a structured ontological
and epistemological architecture.
This paper extracts its formal core
and presents it as an axiomatic system.

\section{Formal Language}

\subsection{Primitive Sets}

\begin{definition}
Let $\mathcal{B}$ denote the set of ontological states (Being).
\end{definition}

\begin{definition}
Let $\mathcal{K}$ denote the set of knowledge states.
\end{definition}

\begin{definition}
Let $\mathcal{E}$ denote the set of ethical calibration functions.
\end{definition}

\section{Axiomatic Core}

\begin{axiom}[Ontological Ground]
There exists a non-empty base state $b_0 \in \mathcal{B}$.
\end{axiom}

\begin{axiom}[Epistemic Mapping]
There exists a mapping
\[
\phi : \mathcal{B} \to \mathcal{K}
\]
such that knowledge states are derived from ontological states.
\end{axiom}

\begin{axiom}[Ethical Calibration]
There exists a function
\[
\psi : \mathcal{K} \to \mathcal{E}
\]
that constrains knowledge application.
\end{axiom}

\section{Consistency Sketch}

\begin{proposition}
If $\mathcal{B} \neq \emptyset$, then
$\mathcal{K}$ and $\mathcal{E}$ are constructible.
\end{proposition}

\begin{proof}
Given Axiom 1, $b_0$ exists.
By Axiom 2, $\phi(b_0)$ exists.
By Axiom 3, $\psi(\phi(b_0))$ exists.
Hence the layered system is non-empty.
\end{proof}

\section{Conclusion}

We defined a minimal consistent formal core
of the Hebûn--Zanabûn system.
Future work includes type-theoretic formalization
and machine verification in Lean or Coq.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
