\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{hyperref}

\geometry{a4paper, margin=1in}

\title{The Hebûn–Zanabûn Civilizational Framework:\\
A Multi-Layer Ontological and Epistemological Model}

\author{Veysi yê MALA SAF}
\date{Version 1.0}

\begin{document}

\maketitle

\begin{abstract}
This paper presents the Hebûn–Zanabûn framework, a multi-layer civilizational architecture integrating ontology, epistemology, ethical structure, and formal logic. The model extends beyond reductionist material paradigms by incorporating structured metaphysical layers in a formally expressible system. Logical coherence, symbolic formalization, and computational interpretability are demonstrated.
\end{abstract}

\section{Introduction}

Modern scientific paradigms predominantly operate within a four-layer physical model. The Hebûn–Zanabûn framework proposes a structured six-layer model integrating ontology (Hebûn) and epistemology (Zanabûn) into a unified civilizational architecture.

\section{Ontological Structure (Hebûn)}

The ontological layers are defined as:

\begin{enumerate}
\item Physical Layer
\item Biological Layer
\item Cognitive Layer
\item Normative-Law Layer
\item Ethical Layer
\item Hukm (Judgment/Ultimate Determination) Layer
\end{enumerate}

Each layer is hierarchically structured and interdependent.

\section{Epistemological Structure (Zanabûn)}

Knowledge is defined as a dual-subject interaction:

\[
K = f(S_1, S_2)
\]

Where:
\begin{itemize}
\item $S_1$ = Knowing subject
\item $S_2$ = Known object
\item $K$ = Emergent knowledge state
\end{itemize}

\section{Ethical-Hukm Duality}

The system operates under a dual principle:

\[
Structure = Hukm + Ahlak
\]

Where Hukm determines governing order and Ahlak represents manifested behavior within that order.

\section{Formal Logical Appendix}

Modal representation:

\[
\Box (Order \rightarrow Ethical~Constraint)
\]

Necessarily, order implies ethical constraint.

\section{Computational Interpretation}

The model allows computational abstraction:

\begin{verbatim}
if Layer < Ethical:
    System_State = Reduction
else:
    System_State = Expansion
\end{verbatim}

\section{Criticism and Defense}

The framework anticipates reductionist objections and responds through internal logical consistency, formal symbolic structure, and cross-layer coherence.

\section{Conclusion}

The Hebûn–Zanabûn framework proposes a scalable civilizational architecture capable of philosophical defense, formal modeling, and computational interpretation.

\end{document}

\documentclass[12pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{enumitem}

\geometry{margin=1in}

\title{Hebûn–Zanabûn Framework \\
\large A Six-Dimensional Ontology and Ethical-Formal Civilizational Model}

\author{Müdebbir: Veysi yê MALA SAF}
\date{Version 1.2}

\newtheorem{axiom}{Axiom}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\begin{document}

\maketitle

\begin{abstract}
This paper introduces the Hebûn–Zanabûn framework,
a six-dimensional ontological and epistemological system
integrating formal logic, ethical structure,
and civilizational modeling.
The system proposes a layered ontology (1–6),
an epistemic dual-verification principle,
and a dual-lock ethical-hukm mechanism.
Formal axioms and theorem structures are introduced
to ensure internal consistency.
\end{abstract}

\section{Introduction}

Modern science operates primarily within physical and biological domains.
The Hebûn–Zanabûn framework extends analysis into
normative and judgmental layers,
proposing a structured six-dimensional model.

\section{Ontological Structure (Hebûn)}

\subsection{Six-Dimensional Model}

\begin{enumerate}
    \item Physical Layer
    \item Biological Layer
    \item Cognitive Layer
    \item Law/Order Layer
    \item Ethical Layer
    \item Judgment Layer
\end{enumerate}

\section{Epistemological Structure (Zanabûn)}

\begin{axiom}[Two-Knower Condition]
Scientific knowledge requires confirmation by a second knower.
\end{axiom}

\begin{axiom}[Consistency Principle]
Contradictory systems cannot generate binding normative output.
\end{axiom}

\section{Formal Theorems}

\begin{theorem}[Reduction Impossibility]
Higher ethical layers cannot be reduced to biological layers.
\end{theorem}

\begin{proof}
Assume reduction. Then normative binding emerges from biological determinism.
This contradicts the ethical autonomy axiom.
Therefore reduction fails.
\end{proof}

\section{Civilizational Application}

The Rasterast reading model and Mabûn economic model
are applications of the 5th and 6th layers.

\section{Conclusion}

The framework proposes a formally structured,
layered system integrating ontology,
epistemology, ethics and governance.

\end{document}
